The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-02-15 src/ms.tex
more text about astroquery

Summary of Astroquery and engagement with archives. Cite astroquery paper.

>>>>

Astroquery \cite{Ginsburg2019} is the astropy-coordinated package for interacting with online archives of astronomical and related data. It contains over 50 modules for querying astronomical databases, large and small. In particular, since \PaperII, significant contributions to \texttt{astroquery} have come from several of the major archives, including the European Space Agency (ESA), the Mikulski Archive for Space Telescopes (MAST) at the Space Telescope Science Institute (STScI), the Infrared Science Archive at the NASA Infrared Processing and Analysis Center (IRSA at NASA IPAC), the Canadian Astronomical Data Center (CADC), and the Atacama Large Millimeter/Submillimeter Array archive (ALMA). These contributions represent a formal acknowledgement of the utility of a centralized tool suite for archive interaction.

The widespread usage of \texttt{astroquery} is apparent from the range of keywords represented in the journal articles which cite \cite{Ginsburg2019}; everything from astroids to galaxies is represented. Additionally Astroquery functionality has been built into several special purpose Python packages, such as \texttt{LightKurve} \cite{LightKurve} and SORA \cite{SORA}]. More than 4,000 repositories on GitHub make use of \texttt{astroquery} in some way, an \texttt{astroquery} features in a large number of tutorials on various facets of astronomical analysis.

Many of the existing and newly-contributed tools rely on Virtual Observatory (VO) tools.
These use the underlying package \texttt{pyvo}, which has also recently become an astropy-coordinated package. While many or all of the functions provided in \texttt{astroquery} can be achieved through direct use of VO tools implemented in \texttt{pyvo}, the \texttt{astroquery} interfaces more closely resemble the web interfaces more familiar to users.
% note to coauthors: we need a section on pyvo
